---
title: "Yelp Me Pls"
author: "Carmelia Muljadi, Madeline Wu, Brenda Zhang"
date: "May 2, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE}
library(DataComputing)
library(dplyr)
library(ggthemes)
library(readr)
library(maps)
```


```{r, include = FALSE}
#import and read csv files into R

business <- read_csv("C:\\Users\\Brenda\\Documents\\UC Berkeley\\Sophomore\\Spring 2017\\STAT 133\\Final Project\\dataset-examples-master\\yelp_academic_dataset_business.csv")

review <- read_csv("C:\\Users\\Brenda\\Documents\\UC Berkeley\\Sophomore\\Spring 2017\\STAT 133\\Final Project\\dataset-examples-master\\yelp_academic_dataset_review.csv")

```

Data Wrangling

```{r, include = FALSE}
# cleaning data to make business_clean and review_clean

business_clean <- business %>%
  select(business_id, city, stars, review_count, is_open, categories, latitude,longitude) %>% 
  filter(is_open == 1) %>% # filtering out businesses that are no longer in business
  select(business_id, city, stars, review_count, categories, latitude, longitude) %>% # selecting relevant columns
  rename(avg_stars = stars) 
  
review_clean <- review %>%
  select(business_id, date, stars) %>% # select relevant columns
  rename(review_stars = stars)

```


```{r, include = FALSE}
# Making separate business data tables for each of the 16 categories

arts <- business_clean %>% filter(grepl("Arts & Entertainment", categories))
nrow(arts)

active <- business_clean %>% filter(grepl("Active Life", categories))
nrow(active)

food <- business_clean %>% filter(grepl("Food|Restaurants", categories))
nrow(food)

beauty <- business_clean %>% filter(grepl("Beauty & Spas", categories))
nrow(beauty)

edu <- business_clean %>% filter(grepl("Education", categories))
nrow(edu)

event <- business_clean %>% filter(grepl("Event Planning & Services", categories))
nrow(event)

fin <- business_clean %>% filter(grepl("Financial Planning & Services", categories))
nrow(fin)

health <- business_clean %>% filter(grepl("Health & Medical", categories))
nrow(health)

home <- business_clean %>% filter(grepl("Home Services", categories))
nrow(home)

hotel <- business_clean %>% filter(grepl("Hotels & Travel", categories))
nrow(hotel)

pets <- business_clean %>% filter(grepl("Pets", categories))
nrow(pets)

prof <- business_clean %>% filter(grepl("Professional Services", categories))
nrow(prof)

publicserv <- business_clean %>% filter(grepl("Public Services & Government", categories))
nrow(publicserv)

realestate <- business_clean %>% filter(grepl("Real Estate", categories))
nrow(realestate)

relig <- business_clean %>% filter(grepl("Religious Organizations", categories))
nrow(relig)

shopping <- business_clean %>% filter(grepl("Shopping", categories))
nrow(shopping)

# Creating a two column data table of categories and their number of associated businesses

categ_count <- data.frame(
  categories = c("Arts", "Active", "Food", "Beauty", "Education", "Events", "Finance", "Health", "Home", "Hotel", "Pets", "Professional", "Public Services", "Real Estate", "Religious", "Shopping"),
  count = c(nrow(arts), nrow(active), nrow(food), nrow(beauty), nrow(edu), nrow(event), nrow(fin), nrow(health), nrow(home), nrow(hotel), nrow(pets), nrow(prof), nrow(publicserv), nrow(realestate), nrow(relig), nrow(shopping))
)

# Create a visualization that plots the count for each category
categ_count_bar <- categ_count %>% ggplot(aes(x = categories,y = count)) + geom_col(aes(fill = categories)) + labs(title = "Number of Businesses in Each Category", x = "Categories", y="Number of Business") + theme(axis.text.x = element_text(angle = 90))

categ_count_bar
```


```{r, include = FALSE}
# Add an additional column with the category name for each table
food1 <- food %>% mutate (categories = "Food")
shopping1 <- shopping %>% mutate (categories = "Shopping")
beauty1 <- beauty %>% mutate (categories = "Beauty")
home1 <- home %>% mutate(categories = "Home")
arts1 <- arts %>% mutate(categories = "Arts")
active1 <- active %>% mutate(categories = "Active")
edu1<- edu %>% mutate(categories = "Education")
event1 <- event %>% mutate(categories = "Events")
fin1 <- fin %>% mutate(categories = "Finance")
health1 <- health %>% mutate(categories = "Health")
hotel1 <- hotel %>% mutate(categories = "Hotel")
pets1 <- pets %>% mutate(categories = "Pets")
prof1 <- prof %>% mutate(categories = "Professional")
publicserv1 <- publicserv %>% mutate(categories = "Public Services")
realestate1 <- realestate %>% mutate(categories = "Real Estate")
relig1 <- relig %>% mutate(categories = "Religious")

# Business table for food, shopping, beauty, and home businesses, with updated category column
business_filtered <- rbind(food1,shopping1,beauty1,home1)

# Business table for all categories, with updated category column
business_all <- rbind(business_filtered, arts1, active1,edu1, event1,fin1,health1,hotel, pets1,prof1,publicserv1,realestate1,relig1)

# Making a table mapping each month to season
seasons <- data.frame(season = c("Winter","Winter","Winter","Spring","Spring","Spring","Summer","Summer","Summer","Fall","Fall","Fall"), month = c(12,1,2,3,4,5,6,7,8,9,10,11))
seasons

# Making table mapping each month number to a month name
months <- data.frame(month = seq(1,12), name = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec"))
months

# Create fundamental data set for filtered categories, associating each review with its business
data <- business_filtered %>%
  left_join(review_clean, by="business_id") %>%
  select(business_id, city, categories, date, review_stars, longitude, latitude, review_count)

# Create fundamental data set for all categories, associating each review with its business

data_all <- business_all %>%
  left_join(review_clean, by="business_id") %>%
  select(business_id, city, categories, date, review_stars)

```

Answering research question 1

```{r, include = FALSE}
# Creates a table mapping each month to the number of reviews for each category

food_joined <- data %>%
  filter(categories == "Food") %>% 
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month")
View(food_joined)

shopping_joined <-data %>%
  filter(categories == "Shopping") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month")

beauty_joined <- data %>%
  filter(categories == "Beauty") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month")

home_joined <- data %>%
  filter(categories == "Home") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month")

```

```{r, include = FALSE}

# Creating individual tables for each category: month, number of reviews, category

food_joined_all <- data_all %>%
  filter(categories == "Food") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Food")

shopping_joined_all <-data_all %>%
  filter(categories == "Shopping") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Shopping")

beauty_joined_all <- data_all %>%
  filter(categories == "Beauty") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Beauty")

home_joined_all <- data_all %>%
  filter(categories == "Home") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Home")

arts_joined_all <- data_all %>%
  filter(categories == "Arts") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Arts")

active_joined_all <- data_all %>%
  filter(categories == "Active") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Active")

edu_joined_all <- data_all %>%
  filter(categories == "Education") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Education")

event_joined_all <- data_all %>%
  filter(categories == "Events") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Events")

fin_joined_all <- data_all %>%
  filter(categories == "Finance") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Finance")

health_joined_all <- data_all %>%
  filter(categories == "Health") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Health")

hotel_joined_all <- data_all %>%
  filter(categories == "Hotel") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Hotel")

pets_joined_all <- data_all %>%
  filter(categories == "Pets") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Pets")

prof_joined_all <- data_all %>%
  filter(categories == "Professional") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Professional")

publicserv_joined_all <- data_all %>%
  filter(categories == "Public Services") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Public Services")

realestate_joined_all <- data_all %>%
  filter(categories == "Real Estate") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Real Estate")

relig_joined_all <- data_all %>%
  filter(categories == "Religious") %>%
  mutate(month = lubridate::month(date)) %>%
  group_by(month) %>%
  summarize(num_reviews = n()) %>% 
  left_join(months, by="month") %>%
  mutate(categories = "Religious")

# Create aggregate table of reviews per business, grouped by category and month

top_categ <- rbind(food_joined_all, shopping_joined_all,beauty_joined_all,home_joined_all,arts_joined_all,active_joined_all, edu_joined_all, event_joined_all, fin_joined_all, health_joined_all, hotel_joined_all, pets_joined_all, prof_joined_all, publicserv_joined_all, realestate_joined_all, relig_joined_all) %>%
  left_join(categ_count, by = "categories") %>% # create an intermediate table with month, number of reviews, number of businesses, category 
  mutate(reviews_per_business = num_reviews/count) %>% # create column with our quantifier
  left_join(seasons, by = "month") %>%
  group_by(categories, season) %>%
  summarise(reviews_per_business = mean(reviews_per_business))

top_categ

```

```{r, include=FALSE}

# Function that calculates percent change
calc_per_change <- function(x) {
  change_vec <- c(0)
  for (i in 1:length(x)-1) {
    change <- (x[i+1]-x[i])/x[i]
    change_vec <-append(change_vec, change)
  }
  return(change_vec)
}

# Creates a table with added column for percent change for each category

food_joined_all_change <- food_joined_all %>% 
  mutate(percent_change = calc_per_change(food_joined_all$num_reviews))

shopping_joined_all_change <- shopping_joined_all %>% 
  mutate(percent_change = calc_per_change(shopping_joined_all$num_reviews))

home_joined_all_change <- home_joined_all %>% 
  mutate(percent_change = calc_per_change(home_joined_all$num_reviews))

beauty_joined_all_change <- beauty_joined_all %>% 
  mutate(percent_change = calc_per_change(beauty_joined_all$num_reviews))

# Creates aggregate precent change table for all selected categories

per_change <- rbind(food_joined_all_change, shopping_joined_all_change, beauty_joined_all_change, home_joined_all_change)

per_change

```

Answering research question 2
```{r, include = FALSE}
# Creates a table mapping each month to the category's average resulting rating during that month

food_joined_ratings <- data %>%
  filter(categories == "Food") %>%
  group_by(date) %>%
  summarize(ave_rating = mean(review_stars)) %>%
  mutate(month = lubridate::month(date)) %>%
  left_join(months, by="month")


shopping_joined_ratings <- data %>%
  filter(categories == "Shopping") %>%
  group_by(date) %>%
  summarize(ave_rating = mean(review_stars)) %>%
  mutate(month = lubridate::month(date)) %>%
  left_join(months, by="month")

beauty_joined_ratings <- data %>%
  filter(categories == "Beauty") %>%
  group_by(date) %>%
  summarize(ave_rating = mean(review_stars)) %>%
  mutate(month = lubridate::month(date)) %>%
  left_join(months, by="month")

home_joined_ratings <- data %>%
  filter(categories == "Home") %>%
  group_by(date) %>%
  summarize(ave_rating = mean(review_stars)) %>%
  mutate(month = lubridate::month(date)) %>%
  left_join(months, by="month")


```

Answering research question 3 
```{r, include=FALSE}
# Create summary table for number of reviews for a given location in each season, limiting data set to certain cities
geo <- data %>%
  mutate(latitude = round(latitude, digits = 0), longitude = round(longitude, digits = 0)) %>% # cleaning coordinates 
  mutate(month = lubridate::month(date)) %>% # transforming month column into glyph-ready data
  left_join(seasons, by = "month") %>% # add a column associating a review with its season
  group_by(longitude, latitude, season) %>% 
  mutate(tot_numrev = sum(review_count)) %>% 
  select(city,tot_numrev, season, latitude, longitude) %>%
  unique() %>%
  filter(city == "New York City"|city == "Las Vegas"|city =="Chicago"|city =="Pittsburgh"|city =="Charlotte"|city =="Phoenix"|city =="San Francisco"|city =="Los Angeles"|city =="Houston"|city =="Boston") # select desired cities
```

## Analyzing Reviewing Behavior In Relation to Seasonality

**Key Questions**

* What businesses are most popular (are reviewed the most) during each season? Which categories are most affected by seasonal trends?  
* For a given business/category, how do reviews change with the seasons? 
* How does reviewing behavior change for a given category across different geographical regions? 

**Data Source**

* Yelp Academic Dataset:
  + yelp academic dataset business.json
  + yelp academic dataset review.json
* Variables of interest: business_id, city, stars, review_count, is_open, categories, date, latitude, longitude
* Time period: 2004-2015

**Methods**

* from this class: lubridate, data verbs for data wrangling, regex and grepl, join, ggplot (geom_col, geom_violin, geom_line), themes
* outside of this class: vertically combining data tables (rbind), converting json to csv


## Results: 

### Number of Businesses in Each Category 

```{r, echo = FALSE}
categ_count_bar <- categ_count %>% ggplot(aes(x = categories,y = count)) + geom_col(aes(fill = categories)) + labs(title = "Number of Businesses in Each Category", x = "Categories", y="Number of Business") + theme(axis.text.x = element_text(angle = 90))

categ_count_bar
```

### Top Categories Per Season

```{r, echo = FALSE}
top_categ_bar <- top_categ %>% ggplot(aes(x=categories, y = reviews_per_business)) + geom_col(aes(fill=categories)) + facet_wrap(~season) + labs(title = "Top Categories for Each Season", x = "Categories", y="Reviews Per Business") + theme(axis.text.x = element_text(angle = 90))

top_categ_bar
```

### Number of Reviews Over Time for Each Category

```{r, echo=FALSE}
food_joined %>% ggplot(aes(x = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y=num_reviews)) + geom_col(fill = "sky blue") + labs(title = "Food",x = "Month", y= "Number of Reviews") + theme(plot.title = element_text(size = 25, color = "black")) + theme_igray()

shopping_joined %>% ggplot(aes(x = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y=num_reviews)) + geom_col(fill = "purple") + labs(title = "Shopping",x = "Month", y= "Number of Reviews") + theme(plot.title = element_text(size = 25, color = "black")) + theme_igray()

beauty_joined %>% ggplot(aes(x = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y=num_reviews)) + geom_col(fill = "pink") + labs(title = "Beauty",x = "Month", y= "Number of Reviews") + theme(plot.title = element_text(size = 25, color = "black")) + theme_igray()

home_joined %>% ggplot(aes(x = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y=num_reviews)) + geom_col(fill = "dark green") + labs(title = "Home",x = "Month", y= "Number of Reviews") + theme(plot.title = element_text(size = 25, color = "black")) + theme_igray()
```

### Percent Change Per Month

```{r, echo = FALSE}
per_change_line <- per_change %>%
  ggplot(aes(x=factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y=percent_change, col = categories, group=categories)) + geom_line() + geom_point() + labs(title = "Percent Change of Reviews For Each Category", x = "Month", y="Percent Change")

per_change_line

```

## Results: 


### Average Rating Over Time for Each Category

```{r, echo = FALSE}
food_joined_ratings %>% ggplot(aes(x = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y = ave_rating)) + geom_violin(fill = "sky blue") + scale_y_continuous(breaks = seq(0,5,0.5)) + labs(title = "Food",x = "Month", y= "Average Rating") + theme(plot.title = element_text(size = 20, color = "black"))

shopping_joined_ratings %>% ggplot(aes(x = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y = ave_rating)) + geom_violin(fill = "purple")+ scale_y_continuous(breaks = seq(0,5,0.5)) + labs(title = "Shopping",x = "Month", y= "Average Rating") + theme(plot.title = element_text(size = 20, color = "black"))

beauty_joined_ratings %>% ggplot(aes(x  = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y = ave_rating)) + geom_violin(fill = "pink")+ scale_y_continuous(breaks = seq(0,5,0.5)) + labs(title = "Beauty",x = "Month", y= "Average Rating") + theme(plot.title = element_text(size = 20, color = "black"))

home_joined_ratings %>% ggplot(aes(x  = factor(name, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov","Dec")), y = ave_rating)) + geom_violin(fill = "dark green")+ scale_y_continuous(breaks = seq(0,5,0.5)) + labs(title = "Home",x = "Month", y= "Average Rating") + theme(plot.title = element_text(size = 20, color = "black"))


```

### Number of Reviews By City Per Season

```{r, echo = FALSE, fig.width=11, fig.height=8}
 
map <-map_data("state")

          
p <- ggplot() + geom_polygon( data=map, aes(x=long, y=lat, group = group),colour="white", fill="grey10" )
p1 <- p + geom_point( data=geo, aes(x=longitude, y=latitude, size = tot_numrev), color="coral1") + scale_size(name="Total Number of Reviews") + facet_wrap(~season) + labs(title = "Number of Reviews By City Per Season", x="Latitude", y="Longitude")

p1


```

